---
title: "sheltering_counties"
author: "David McCoy"
date: "6/27/2020"
output: html_document
---




```{r}
library(ggplot2)
library(COVID19)
library(dplyr)
library(zoo)
library(ggplot2)
library(viridis)
library(glue)
library(tools)
library(lubridate)
library(gifski)
library(gganimate)
library(summarytools)
library(RCurl)
library(forcats)

```

This code creates an animated gif file that shows the relationship between the public's mobility and reported deaths at the county level thirty days later. The values are 7-day rolling averages. The data includes all counties with 390,000 residents. This cutoff was chosen to include Orleans Parrish in the data.  

The sources include [Google's Mobility Data](https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv), The [NYT's public repository of data](https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv), and the [COVID19](https://github.com/datasets/covid-19) API for R.  



# Counties Cases, Deaths, Mobility: Get data
```{r}
gmob <- "https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv"
gmob <- getURL(gmob)
gmob <- read.csv(textConnection(gmob))
#saveRDS(gmob, "data/backup/Global_Mobility_Report_june26.rds")

counties <- "https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv"
counties <- getURL(counties)
counties <- read.csv(textConnection(counties))
#saveRDS(counties, "data/backup/nyt_counties_june26.rds")

cc <- covid19("US", level=3, cache=T)
#saveRDS(cc, "data/backup/covid19api_counties_june26.rds")
```

# Clean Data
```{r}
cc <- readRDS("data/backup/covid19api_counties_june26.rds")
counties <- readRDS("data/backup/nyt_counties_june26.Rds")
gmob <- readRDS("data/backup/Global_Mobility_Report_june26.rds")

# Extract time invariant data
id <- cc %>%
  ungroup() %>%
  mutate(
    id = key_numeric) %>%
  group_by(id) %>%
  summarize(
    population = population[1],
    fips = id[1],
    county = administrative_area_level_3[1],
    state = administrative_area_level_2[1]) %>%
  ungroup %>%
  select(-id)

# Clean COVID19 API data
c <- cc %>%
  ungroup() %>%
  mutate(
    fips = key_numeric,
    deaths2 = deaths,
    cases2 = confirmed) %>%
  select(fips, date, cases2, deaths2) %>%
  arrange(fips, date)

# Clean NYT data
nyt <- counties %>%
  mutate(
    date = as.Date(date)) %>%
  select(-state, -county)

# Mobility at county - Many missing values AND counties but large look OK
m <- gmob %>%
  filter(country_region_code=="US" & census_fips_code != "" & iso_3166_2_code=="") %>%
  mutate(
    date = as.Date(date),
    fips = census_fips_code,
    Recreation = retail_and_recreation_percent_change_from_baseline,
    Grocery = grocery_and_pharmacy_percent_change_from_baseline,
    Parks = parks_percent_change_from_baseline,
    Transit = transit_stations_percent_change_from_baseline,
    Work = workplaces_percent_change_from_baseline,
    n.Home = - residential_percent_change_from_baseline) %>%
  select(fips, date, Recreation, Grocery, Parks, Transit, Work, n.Home)

# Most inclusive covid and mobility data
d <- full_join(c, nyt, by=c("fips", "date"))
d <- full_join(d, m, by=c("fips", "date"))
d <- full_join(d, id, by="fips")

rescale = 100000

shelter_county <- d %>%
  filter(population>390000 & !county %in% c("Kansas City", "New York City", "District of Columbia")) %>%
  arrange(fips, date) %>%
  group_by(fips) %>%
  mutate(
    cases = ifelse(is.na(cases), 0, cases),
    deaths = ifelse(is.na(deaths), 0, deaths),
    dailydeaths = deaths - lag(deaths),
    dailydeaths  = ifelse(dailydeaths<0 , 0, dailydeaths),
    dailycases = cases - lag(cases),
    dailycases = ifelse(dailycases<0 , 0, dailycases),
    dailydeaths.pc = (dailydeaths/population) * rescale,
    dailycases.pc = (dailycases/population) * rescale,
    r.dailydeaths = rollmean(dailydeaths, k = 7, fill = NA),
    r.dailydeaths.log = log(1+(rollmean(dailydeaths, k = 7, fill = NA))),
    r.dailydeaths.pc = rollmean(dailydeaths.pc, k = 7, fill = NA),
    r.dailycases = rollmean(dailycases, k = 7, fill = NA),
    r.dailycases.log = log(1+(rollmean(dailycases, k = 7, fill = NA))),
    r.dailycases.pc = rollmean(dailycases.pc, k = 7, fill = NA),
    Shelter.old = - (Recreation + n.Home + Work + Transit)/4,
    Shelter = - rowMeans(cbind(Recreation, n.Home, Work, Transit), na.rm=T),
    r.Shelter = rollmean(Shelter, k=7, fill=NA),
    Date = format(as.Date(date), "%b %d"),
    Date.L30 = lag(Date, 30L),
    Shelter.L30 = lag(Shelter, 30L),
    r.Shelter.L30 = lag(r.Shelter, 30L),
    r.dailycases.pc.L30 = 1 + lag(r.dailycases.pc, 30L)) %>%
  ungroup() %>%
  filter(date > as.Date("2020-03-18") & date < as.Date("2020-06-20")) %>%
  group_by(fips) %>%
  mutate(n = 1:n()) %>%
  ungroup() 

# Order Date by n
shelter_county$Date <- fct_reorder(shelter_county$Date, shelter_county$n)
shelter_county$Date.L30 <- fct_reorder(shelter_county$Date.L30, shelter_county$n)

rm(c, cc, counties, d, id, m, gmob, nyt)
```

# Plot gganimate: Deaths
```{r}
p <- ggplot(data = shelter_county,
            aes(x=r.Shelter.L30, 
                y=r.dailydeaths.pc, 
                color=r.dailycases.pc.L30)) +
  geom_point(size=4) + 
  #xlim(-40, 62) + ylim(0, 8) +
  scale_color_viridis(option="plasma", alpha = .2, 
                      limits=c(1,115), trans = "log", 
                      breaks = c(1, 5, 20, 110)) +
  #theme(text=element_text(family="serif")) +
  labs(
    title = "Sheltering Behavior and Daily Deaths per 100k Residents \n 30 Days Later: All US Counties with 400K Residents \n {closest_state}",
    subtitle = "All values are 7-Day Rolling Averages",
    caption = "Data Source: Google Mobility Data, NYT, COVID19 API \n Github: DavidRMcCoy",
    color = "Daily Cases",
    y = "Daily Deaths per 100k Residents: 30 Days Later",
    x = "Google Mobility Index: Staying Home + Working (rev.) + Recreation (rev.)") +
  transition_states(Date.L30,
                    transition_length = 10,
                    state_length = 1,
                    wrap = F) +
  shadow_wake(wake_length = 0.005) + enter_fade() + exit_shrink()


c_s <- animate(p, nframes = 186, detail=20)

anim_save("county_sheltering.gif", animation=c_s)

```





